I would first thank Emery Berger, my Ph.D. thesis advisor, for his wonderful supervision and enthusiastic support in the development of this thesis work. Emery taught me to work on those important and practical projects, and take everything seriously. I hope that I could be as lively, enthusiastic, and energetic as Emery one day to my students. I also thank my dissertation committee: Scott Kaplan, Yuriy Brun, and Israel Koren, for their valuable insights, feedback, and support. 

I am also fortunate to work with Chen Tian, Timothy Richards, Prashant Shenoy, Ziang Hu, Daniel Waddington, Seetharami Seelam, Wei Tan, Liana Fong, and Arun Iyengar. They provided me valuable guidance and suggestions on those projects we worked on together. I also thank Leeanne Leclerc, James Allan, and Laurie Downey for their help to make my experience at UMASS go as smoothly as possible. 

I couldn't finish my thesis without the help of PLASMA  labmates, including Charlie Curtsinger, Ting Yang, Gene Nowark, Kaituo Li, John Altidor, Divya Krishnan, Dan Barowy, Dimitar Gochev, John Vilk, Nitin Gupta, Jacob Evans, Justin Aquadro, and Emma Tosch. I also feel lucky to meet many friends in the computer science department, including Rui Wang, Ming Li, Tingxin Yan, Zongfang Lin, Kun Tu, Xiaojian Wu, Pengyu Zhang, Bo Jiang, Xiaozhen Tie, Fangyuan Zhou, Hong Zhang, Yue Wang, Wenzhao Liu, etc. I will never forget your help in my life. 

Finally, I would give my special thanks for my wife, Yuyu Tang. She quit her job to support my crazy idea of getting a Ph.D. degree. She also took care of most of the household duties and spent much of her time taking care of our two adorable kids, Yanbin Liu (Grace) and Yanlin Liu (Eileen). My buddy, Guangming Zeng, also deserves my special thanks for his generous help and valuable discussion when I chose the career in the computer science field. I also want to thank my kids, my late grandma, my parents, sisters, and brothers for their understanding and support.    

